### 2023-09-02 sections/2-methods.tex
references fixes and edits to rnn complexity

Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
(i.e. arises from the same generating phenomenon) as the rest of the data. We also rely on the same technical assumption about the self-similarity 
of the dataset made in \cite{yu2015entropy}: if call the number of points within a given 
radius of a point the \emph{density} around that point, we assume that for this radius, the densities around all points
are bounded within a constant multiplicative factor $\gamma$ of each other. Formally, 
we assume that for any point $p$, query $q$, and radius 
$\rho$, \begin{equation} \frac{1}{\gamma}|B(p, \rho)| \leq \mathbb{E}[|B(q, \rho)|] \leq \gamma |B(p, \rho)|. \label{3} \end{equation}

@@

In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search is $\mathcal{O}(\log{2}m)$, 
where $m$ is the metric entropy of the dataset. 
To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
of $\rho$-NN needed to find at least $k$ neighbors.

Because we assume that the query point is sampled from 
the same distribution as the rest of the data, we can assume that the local fractal dimension around the query
is not significantly different from the mean local fractal dimension of the dataset. If the local fractal dimension 
were perfectly uniform throughout the entire dataset, then 
the use of the factor from Equation \ref{2} would guarantee that it only takes 
one iteration of $\rho$-NN to find $k$ points. In practice, [is this the 
way I want to be going with this??]

@@

~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$. By definition of 
$r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.

It remains to provide an estimate for $r'$. As described in \cite{yu2015entropy} bounded density assumption suggests that increase in number of
points reached by doubling the radius has to be roughly uniform across all regions of the manifold. Thus, the global 
average local fractal dimension over all regions of the dataset does not differ significantly from the local fractal dimension around any particular point.
We rely on this fact in our estimation of $r'$.

We let $d'$ be the average of the local fractal dimension across all clusters in the dataset and based on the bounded density assumption, we have that
$d' \approx d_{0}$, where $d_0$ is the local fractal dimension of the root cluster (i.e., $d_0$ is the logarithm of the ratio between the cardinality of the root cluster 
and the cardinality of the ball centered at the root cluster). In this algorithm, we perform $\rho$-nearest neighbors search at equal to the radius of the cluster tree divided by
er's center with half its radius. While, ordinarily we compute local fractal dimension 
by comparing cardinalities of balls centered at some point at two different radii, in order to estimate $r'$, instead compare the cardinality of 
a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of the root cluster} at its radius.

If $C_0$ is the root cluster, $r_0$ is its radius, then we have 
\begin{equation} d_0 = \frac{\log{}\frac{|C_0|}{k}}{\log{}\frac{r_0}{r'}}. \label{4} \end{equation}

Since $d_0$, $|C_0|$, $r_0$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
\begin{equation} r' = r_0\left(\frac{k}{|C_0|}\right)^{\frac{1}{d_0}}. \label{6} \end{equation}

>>>>

Our complexity bounds for $k$-NN by Repeated $\rho$-NN rely on the assumption that the query point is sampled from the same distribution
(i.e. arises from the same generating phenomenon) as the rest of the data. From this assumption, we can infer that the local fractal dimension
at the query point is not significantly different from the mean local fractal dimension of clusters nearby the query.

@@

In ~\cite{ishaq2019clustered}, we claimed that the complexity of coarse search for $\rho$-NN is $\mathcal{O}(\log{2}m)$, 
where $m$ is the metric entropy of the dataset. 
To adjust this bound for $k$-NN by Repeated $\rho$-NN, we must estimate the number of iterations 
of $\rho$-NN needed to find at least $k$ neighbors.

Based on the assumption that the local fractal dimension at the query point is not significantly different than that of 
nearby clusters, the factor from Equation \ref{2} suggests that in the expected case, we need only two iterations of 
$\rho$-NN to find $k$ neighbors: one iteration to find at least one neighbor, and the one more to find enough remaining neighbors.
Since this is a constant factor, it does not affect the asymptotic complexity of coarse search, and thus we have that 
the complexity of coarse search for $k$-NN by Repeated $\rho$-NN is $\mathcal{O}(\log_2{m})$.

@@

~\cite{yu2015entropy} showed that $|F| \leq \gamma |B(q, r')|\left(\frac{r'+ 2r_c}{r'}\right)^d$, where $\gamma$ is a constant. 
By definition of 
$r'$, we have that $|B(q, r')| = k$. Thus, we have that $|F| \leq \gamma k\left(\frac{r'+ 2r_c}{r'}\right)^d$.

It remains to provide an estimate for $r'$. 
To do this, we once again rely on the assumption that the query is from the same distribution as the rest of the data, 
and thus the local fractal dimension at the query point is not significantly different than that of nearby clusters.

We let $d_i$ be the local fractal dimension of some cluster $C_i$ near the query point (i.e., some cluster identified during coarse search).
While ordinarily we compute local fractal dimension 
by comparing cardinalities of two balls centered at \emph{the same} point at two different radii, in order to estimate $r'$, 
we instead compare the cardinality of a ball \emph{around the query} of radius $r'$ to the cardinality \emph{of cluster $C_i$} at its radius $r_i$.
We justify this approach by noting that, since the query is from the same distribution as the rest of the data, we could "drag" the query to 
the center of $C_i$ without significantly changing the local fractal dimension at the query point.

By ~\ref{2}, we then have that \begin{equation} d_i = \frac{\log{}\frac{|C_i|}{k}}{\log{}\frac{r_i}{r'}}. \label{4} \end{equation}

Since $d_i$, $|C_i|$, $r_i$, and $k$ are all known values, we can solve equation ~\ref{4} for $r'$ to get
\begin{equation} r' = r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}. \label{5} \end{equation}

Putting our bounds for coarse and fine search together, we have that $k$-NN by Repeated $\rho$-NN has 
$\mathcal{O}\left(\log_2{m} + k\left(\frac{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}+ 2r_c}{r_i\left(\frac{k}{|C_i|}\right)^{\frac{1}{d_i}}}\right)^d\right)$ 
time complexity, where $m$ is the metric entropy of the dataset, $d$ is the dimensionality of the dataset, $r_i$ is the radius of some cluster $C_i$
near the query ball, $d_i$ is the local fractal dimension of $C_i$, $r_c$ is the radius of leaf clusters, and $k$ is the number of nearest neighbors.

### 2023-09-02 sections/3-datasets.tex
removed quicksort partition
references fixes and edits to rnn complexity

\subsection{APOGEE-2}
\label{subsec:datasets:apogee-2}
The Sloan Digital Sky Survey (SDSS) [3] Apache Point Observatory Galaxy Evolution
Experiment (APOGEE) [4] contains near-infrared
spectra of approximately 130,000 stellar objects in the Milky
Way galaxy. Each spectral datum describes a star's flux in
Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. 
Thus, each datum is a real-valued vector in $\mathbb{R}_{+}^{8575}$.

@@

\subsection{MaNGA}
\label{subsec:datasets:manga}
SDSS's Mapping Nearby Galaxies at APO (MaNGA) [CITE] contains spectral measurements across the face of each of 
about 10,000 galaxies across 2,700 deg$^2$.

@@

TODO: Add details about why this dataset is a good candidate for CAKES.

>>>>

\subsection{APOGEE-2}
\label{subsec:datasets:apogee-2}
The Sloan Digital Sky Survey (SDSS)~\cite{blanton2017sdss} Apache Point Observatory Galaxy Evolution
Experiment (APOGEE)~\cite{alam2015eleventh} contains near-infrared
spectra of approximately 130,000 stellar objects in the Milky
Way galaxy. Each spectral datum describes a star's flux in
Janskys ($1 Jy = 10^{-26} W \dot m^{-2} \dot Hz^{-1}$) at each of 8,575 wavelengths ranging from $\lambda$ = 1.51 µm to 1.70 µm. 
Thus, each datum is a real-valued vector in $\mathbb{R}_{+}^{8575}$.

@@

\subsection{MaNGA}
\label{subsec:datasets:manga}
SDSS's Mapping Nearby Galaxies at APO (MaNGA)~\cite{Talbot2022SDSSIVMA} contains spectral measurements across the face of each of 
about 10,000 galaxies across 2,700 deg$^2$.

@@
